---
title: "COBP Parameters"
author: "Tyson Wepprich"
date: "March 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape)
library(dplyr)
library(ggplot2)
library(popbio)
library(tidyr)
library(stringr)
'%!in%' <- function(x,y)!('%in%'(x,y))
theme_set(theme_bw(base_size = 10))

```

## Intro

As a first step, I'm bringing together demographic data from different studies.


```{r flower counts}
census <- read.csv("cobp.count.csv", header = TRUE)

ggplot(census, aes(x = year, y = flower.count, group = segment)) +
  geom_point() +
    geom_line() +
    facet_wrap(~segment, scales = "free")

```

Floyd and Ranker demography

```{r floyd demography}
data <- read.csv("oldR/matrix.data", header = TRUE)


expanded<-reshape::untable(data[, c(2:5)], data[,1])
expanded$stage <- factor(expanded$stage, levels = c("seed", "small", "medium", "large1", "large2", "flower", "dead"), ordered = TRUE)
expanded$fate <- factor(expanded$fate, levels = c("seed", "small", "medium", "large1", "large2", "flower", "dead"), ordered = TRUE)

stages <- c("seed", "small", "medium", "large1", "large2", "flower")
fates <- c("dead", "small", "medium", "large1", "large2", "flower")

tf <- table(expanded[, "fate"], expanded[, "stage"])
T.mat <- prop.table(tf, 2)


all.mats <- array(data = NA, dim = c(9, 2, 7, 7))
for (plot in 1:9){
  for (year in 1992:1993){
    temp <- expanded[expanded$plot == plot & expanded$year.start == year, ]
    tftmp <- table(temp[, "fate"], temp[, "stage"])
    year.index <- year - 1991
    all.mats[plot, year.index, ,] <- prop.table(tftmp, 2)
  }
}

 temp <- expanded[expanded$plot %in% c(1:3) & expanded$year.start == 1992, ]
    tftmp <- table(temp[, "fate"], temp[, "stage"])
    year.index <- year - 1991
    all.mats[plot, year.index, ,] <- prop.table(tftmp, 2)

```

From Floyd data, the number of flowers in the previous year (n = 18) or 2-years prior (n = 9)
has very little correlation with the number of new seedlings. Seed bank overwhelming
new inputs from flowers.

```{r seedlings from flowers}

oneyr <- data %>% 
  group_by(plot, year.start) %>% 
  summarise(seedlings = sum(num.plants[stage == "seed"]),
                   flowers = sum(num.plants[stage == "flower"]),
                   total.plants = sum(num.plants))

pairs(oneyr[, 3:5])

twoyr <- oneyr %>% 
  group_by(plot) %>% 
  summarise(seedlings = seedlings[2],
            flowers = flowers[1],
            total1 = total.plants[1],
            total2 = total.plants[2])
pairs(twoyr[, 2:5])

# rosette ratio
r2f <- data %>%
  group_by(plot, year.start) %>% 
  summarise(f1 = sum(num.plants[stage == "flower"]),
            f2 = sum(num.plants[fate == "flower"]),
            r1 = sum(num.plants[stage %in% c("small", "medium", "large1", "large2")]),
            r2 = sum(num.plants[fate %in% c("small", "medium", "large1", "large2")])) %>% 
  group_by(plot) %>% 
  summarise(r2f_1992 = r1[1] / f1[1],
            r2f_1993 = r2[1] / f2[1],
            r2f_1994 = r2[2] / f2[2])

```

Other studies with flower capsule to seed data.
Burgess (2005) germination study estimates 1.0, 1.4, and 1.9 viable seeds per capsule (at 4, 16, and 51 months storage).
Munk (2002) estimates number of capsules per flower.
No difference between creeks or treatment on capsules. Clearing plants increases rosettes. Number of capsules not correlated with number of rosettes (similar to F & R).

Roughly 36 viable seeds per flower (1.4 * 26)
  
```{r }

capsules <- c(16.0, 33.7, 28.2, 34.8, 50.8, 22.9, 48.9, 16.1, 33.5, 6.0, 22.4, 0)

```


